---
title: "Delay Data Exploration for Categorical Predictors"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)

library(tidyverse)
library(plotly)

raw_df = read_csv("Reg_data_exp/raw_for_exp.csv")
```

<br>

**We would like to conduct a linear regression on delay time in minutes. In this case, we would need to check if the interactions between some categorical predictors including times of the day, months, and airlines, are significantly associated with the total delay time in minutes. To achieve it, we did some visualizations. For the tidiness of visualizations, we have adjusted the range of axis.**

<br>

Categorical predictors are:

*   **Times of the Day** (no scheduled hour for flight falls within 0:00 - 4:59 in the raw dataset)

    *   Morning: if scheduled hour for flight falls within 5:00 - 8:59 
  
    *   Noon: if scheduled hour for flight falls within 9:00 - 13:59
  
    *   Afternoon: if scheduled hour for flight falls within 14:00 - 17:59
  
    *   Night: if scheduled hour for flight falls within 18:00 - 23:59
    
*   **Months**

    *   Nov: November, 2021
    
    *   Dec: December, 2021
    
    *   Jan: January, 2022

*   **Airlines**

    *   Alaska Airlines
    
    *   American Airlines
    
    *   Delta Air Lines
    
    *   Endeavor Air
    
    *   JetBlue Airways
    
    *   Republic Airways
    
    *   United Air Lines

<br>

```{r}
hour_air = function(air){
  raw_df %>% 
  filter(
    airline == air) %>%     
  mutate(
    mean = mean(delay),
    hour_c = fct_reorder(hour_c, delay)
    ) %>% 
  plot_ly(x = ~hour_c, y = ~delay, color = ~hour_c,
          type = "box", mode = "markers", alpha = .5) %>% 
    layout(
      xaxis = list(title = "Time of the Day"),
      yaxis = list(title = "Delay Time (minutes)", range = c(0, 180)))
}

month_air = function(air){
  raw_df %>% 
  filter(
    airline == air) %>%     
  mutate(
    month = fct_reorder(month, delay)
    ) %>% 
  plot_ly(x = ~month, y = ~delay, color = ~month,
          type = "box", mode = "markers", alpha = .5) %>% 
    layout(
      xaxis = list(title = "Month"),
      yaxis = list(title = "Delay Time (minutes)", range = c(0, 180)))
}

hour_month = function(mon){
raw_df %>% 
  filter(
    month == mon) %>%     
  mutate(
    hour_c = fct_reorder(hour_c, delay)
    ) %>% 
  plot_ly(x = ~hour_c, y = ~delay, color = ~hour_c,
          type = "box", mode = "markers", alpha = .5) %>% 
    layout(
      xaxis = list(title = "Time of the Day"),
      yaxis = list(title = "Delay Time (minutes)", range = c(0, 180)))
}

hour_month_air = function(air){
  raw_df %>% 
    filter(
      airline == air) %>%
    mutate(
      month = fct_reorder(month, date)) %>% 
    plot_ly(x = ~hour_c, y = ~delay, color = ~month,
            type = "box", mode = "markers", alpha = .5) %>% 
    layout(
      boxmode = "group",
      xaxis = list(title = "Time of the Day"),
      yaxis = list(title = "Delay Time (minutes)", range = c(0, 180)))
}

```


## Interaction between Categorical Predictors {.tabset}

### Time*Airline 

#### Before Stratification by Airline

```{r, echo = FALSE}
raw_df %>% 
  mutate(
    mean = mean(delay),
    hour_c = fct_reorder(hour_c, delay)
    ) %>% 
  plot_ly(x = ~hour_c, y = ~delay, color = ~hour_c,
          type = "box", mode = "markers", alpha = .5) %>% 
    layout(
      xaxis = list(title = "Time of the Day"),
      yaxis = list(title = "Delay Time (minutes)", range = c(0, 180)))
```


#### Stratification by Airline {.tabset}

##### Alaska

```{r, echo = FALSE}
hour_air("Alaska Airlines")
```

##### American

```{r, echo = FALSE}
hour_air("American Airlines")
```

##### Delta

```{r, echo = FALSE}
hour_air("Delta Air Lines")
```

##### Endeavor

```{r, echo = FALSE}
hour_air("Endeavor Air")
```

##### JetBlue

```{r, echo = FALSE}
hour_air("JetBlue Airways")
```

##### Republic

```{r, echo = FALSE}
hour_air("Republic Airways")
```

##### United

```{r, echo = FALSE}
hour_air("United Air Lines")
```


### Month*Airline

#### Before Stratification by Airline

```{r, echo = FALSE}
raw_df %>% 
  mutate(
    month = fct_reorder(month, delay)
    ) %>% 
  plot_ly(x = ~month, y = ~delay, color = ~month,
          type = "box", mode = "markers", alpha = .5) %>% 
    layout(
      xaxis = list(title = "Month"),
      yaxis = list(title = "Delay Time (minutes)", range = c(0, 180)))

```


#### Stratification by Airline {.tabset}

##### Alaska

```{r, echo = FALSE}
month_air("Alaska Airlines")
```

##### American

```{r, echo = FALSE}
month_air("American Airlines")
```

##### Delta

```{r, echo = FALSE}
month_air("Delta Air Lines")
```

##### Endeavor

```{r, echo = FALSE}
month_air("Endeavor Air")
```

##### JetBlue

```{r, echo = FALSE}
month_air("JetBlue Airways")
```

##### Republic

```{r, echo = FALSE}
month_air("Republic Airways")
```

##### United

```{r, echo = FALSE}
month_air("United Air Lines")
```


### Month*Time

#### Before Stratification by Month

```{r, echo = FALSE}
raw_df %>% 
  mutate(
    hour_c = fct_reorder(hour_c, delay)
    ) %>% 
  plot_ly(x = ~hour_c, y = ~delay, color = ~hour_c,
          type = "box", mode = "markers", alpha = .5) %>% 
    layout(
      xaxis = list(title = "Time of the Day"),
      yaxis = list(title = "Delay Time (minutes)", range = c(0, 180)))

```

#### Stratification by Month {.tabset}

##### Nov

```{r, echo = FALSE}
hour_month("Nov")
```

##### Dec

```{r, echo = FALSE}
hour_month("Dec")
```

##### Jan

```{r, echo = FALSE}
hour_month("Jan")
```

### Three-ways Interaction

#### Before Stratification by Airline

```{r, echo = FALSE}
raw_df %>% 
  mutate(
    month = fct_reorder(month, date)) %>% 
  plot_ly(x = ~hour_c, y = ~delay, color = ~month,
          type = "box", mode = "markers", alpha = .5) %>% 
    layout(
      boxmode = "group",
      xaxis = list(title = "Time of the Day"),
      yaxis = list(title = "Delay Time (minutes)", range = c(0, 180)))
```

#### Stratification by Airline {.tabset}

##### Alaska

```{r, echo = FALSE}
hour_month_air("Alaska Airlines")
```

##### American

```{r, echo = FALSE}
hour_month_air("American Airlines")
```

##### Delta

```{r, echo = FALSE}
hour_month_air("Delta Air Lines")
```

##### Endeavor

```{r, echo = FALSE}
hour_month_air("Endeavor Air")
```

##### JetBlue

```{r, echo = FALSE}
hour_month_air("JetBlue Airways")
```

##### Republic

```{r, echo = FALSE}
hour_month_air("Republic Airways")
```

##### United

```{r, echo = FALSE}
hour_month_air("United Air Lines")
```

<br>

## Interpretation

Based on the graphs, we observed that between-group differences existed, and adding interaction terms between the categorical predictors could be one of the options for building the linear regression model.

We found that there could be a significant interaction between:

*   Month * Airline
