---
title: "Delay Data Exploration for Categorical Predictors"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)

library(tidyverse)
library(plotly)

raw_df = read_csv("Reg_data_exp/raw_for_exp.csv")
```

<br>


## Interaction between Categorical Predictors {.tabset}
```{r, include = FALSE}
hour_air = function(air){
  raw_df %>% 
  filter(
    airline == air) %>%     
  mutate(
    mean = mean(delay),
    hour_c = fct_reorder(hour_c, delay)
    ) %>% 
  plot_ly(x = ~hour_c, y = ~delay, color = ~hour_c,
          type = 'box', mode = 'markers', alpha = .5) %>% 
    layout(
      xaxis = list(title = "Hour"),
      yaxis = list(title = "Delay Time (minutes)"))
}

month_air = function(air){
  raw_df %>% 
  filter(
    airline == air) %>%     
  mutate(
    month = fct_reorder(month, delay)
    ) %>% 
  plot_ly(x = ~month, y = ~delay, color = ~month,
          type = 'box', mode = 'markers', alpha = .5) %>% 
    layout(
      xaxis = list(title = "Month"),
      yaxis = list(title = "Delay Time (minutes)"))
}

hour_month = function(mon){
raw_df %>% 
  filter(
    month == mon) %>%     
  mutate(
    hour_c = fct_reorder(hour_c, delay)
    ) %>% 
  plot_ly(x = ~hour_c, y = ~delay, color = ~hour_c,
          type = 'box', mode = 'markers', alpha = .5) %>% 
    layout(
      xaxis = list(title = "Hour"),
      yaxis = list(title = "Delay Time (minutes)"))
}

hour_month_air = function(air){
  raw_df %>% 
    filter(
      airline == air) %>%
    mutate(
      month = fct_reorder(month, date)) %>% 
    plot_ly(x = ~hour_c, y = ~delay, color = ~month,
            type = 'box', mode = 'markers', alpha = .5) %>% 
    layout(
      boxmode = "group",
      xaxis = list(title = "Hour"),
      yaxis = list(title = "Delay Time (minutes)"))
}

```

<br>

### Hour*Airline 

#### Before Stratification by Airline
```{r, echo = FALSE}
raw_df %>% 
  mutate(
    mean = mean(delay),
    hour_c = fct_reorder(hour_c, delay)
    ) %>% 
  plot_ly(x = ~hour_c, y = ~delay, color = ~hour_c,
          type = 'box', mode = 'markers', alpha = .5) %>% 
    layout(
      xaxis = list(title = "Hour"),
      yaxis = list(title = "Delay Time (minutes)"))
```

#### Stratification by Airline {.tabset}
##### Alaska
```{r, echo = FALSE}
hour_air("Alaska Airlines")
```

##### American
```{r, echo = FALSE}
hour_air("American Airlines")
```

##### Delta
```{r, echo = FALSE}
hour_air("Delta Air Lines")
```

##### Endeavor
```{r, echo = FALSE}
hour_air("Endeavor Air")
```

##### JetBlue {.active}
```{r, echo = FALSE}
hour_air("JetBlue Airways")
```

##### Republic
```{r, echo = FALSE}
hour_air("Republic Airways")
```

##### United
```{r, echo = FALSE}
hour_air("United Air Lines")
```


### Month*Airline
#### Before Stratification by Airline
```{r, echo = FALSE}
raw_df %>% 
  mutate(
    month = fct_reorder(month, delay)
    ) %>% 
  plot_ly(x = ~month, y = ~delay, color = ~month,
          type = 'box', mode = 'markers', alpha = .5) %>% 
    layout(
      xaxis = list(title = "Month"),
      yaxis = list(title = "Delay Time (minutes)"))

```


#### Stratification by Airline {.tabset}
##### Alaska
```{r, echo = FALSE}
month_air("Alaska Airlines")
```

##### American
```{r, echo = FALSE}
month_air("American Airlines")
```

##### Delta
```{r, echo = FALSE}
month_air("Delta Air Lines")
```

##### Endeavor
```{r, echo = FALSE}
month_air("Endeavor Air")
```

##### JetBlue {.active}
```{r, echo = FALSE}
month_air("JetBlue Airways")
```

##### Republic
```{r, echo = FALSE}
month_air("Republic Airways")
```

##### United
```{r, echo = FALSE}
month_air("United Air Lines")
```


### Month*Hour
#### Before Stratification by Month
```{r, echo = FALSE}
raw_df %>% 
  mutate(
    hour_c = fct_reorder(hour_c, delay)
    ) %>% 
  plot_ly(x = ~hour_c, y = ~delay, color = ~hour_c,
          type = 'box', mode = 'markers', alpha = .5) %>% 
    layout(
      xaxis = list(title = "Hour"),
      yaxis = list(title = "Delay Time (minutes)"))

```

#### Stratification by Month {.tabset}
##### Nov
```{r, echo = FALSE}
hour_month("Nov")
```

##### Dec
```{r, echo = FALSE}
hour_month("Dec")
```

##### Jan
```{r, echo = FALSE}
hour_month("Jan")
```

### Three-ways Interaction
#### Before Stratification by Airline
```{r, echo = FALSE}
raw_df %>% 
  mutate(
    month = fct_reorder(month, date)) %>% 
  plot_ly(x = ~hour_c, y = ~delay, color = ~month,
          type = 'box', mode = 'markers', alpha = .5) %>% 
    layout(
      boxmode = "group",
      xaxis = list(title = "Hour"),
      yaxis = list(title = "Delay Time (minutes)"))
```

#### Stratification by Airline {.tabset}
##### Alaska
```{r, echo = FALSE}
hour_month_air("Alaska Airlines")
```

##### American
```{r, echo = FALSE}
hour_month_air("American Airlines")
```

##### Delta
```{r, echo = FALSE}
hour_month_air("Delta Air Lines")
```

##### Endeavor
```{r, echo = FALSE}
hour_month_air("Endeavor Air")
```

##### JetBlue {.active}
```{r, echo = FALSE}
hour_month_air("JetBlue Airways")
```

##### Republic
```{r, echo = FALSE}
hour_month_air("Republic Airways")
```

##### United
```{r, echo = FALSE}
hour_month_air("United Air Lines")
```


