---
title: "Could Destination Airport Affect Delay and Cancellation?"
output:   
  html_document:
    toc: true
    toc_float: true

---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)

library(tidyverse)
library(plotly)


delay = read_csv("tidied_data/delay.csv")
cancel = read_csv("tidied_data/cancel.csv")

```

<br>


**There are `r n_distinct(delay$destination_airport)` and `r n_distinct(cancel$destination_airport)` destination airports in the delay and cancellation datasets, respectively. For the efficiency purpose in statistical analysis, we do not include destination airport as a predictor in our models.** <br>
**However, it is still an interesting factor to explore.**

<br>


## Overall Trend 


First we would like to check if delay and cancellation counts differ in different destination airports.

<br>
```{r, echo = F}
delay %>% 
  group_by(destination_airport) %>% 
  summarize(
    count = n()) %>% 
  mutate(
    destination_airport = fct_reorder(destination_airport, count),
    text_label = str_c("Airport: ", destination_airport, "\nCount: ", count)
      ) %>% 
  plot_ly(x = ~destination_airport, y = ~count, text = ~text_label, hoverinfo = "text",
          color = "rgb(255, 65, 54)", type = "bar", alpha = .5) %>% 
  layout(
    xaxis = list(title = "Destination Airport"),
    yaxis = list(title = "Count"),
    title = "Distribution of Delay by Destination Airport")
```

Flights from JFK to `r delay %>% count(destination_airport) %>% slice(which.max(n)) %>% pull(destination_airport)` have the highest delay occurrences with a number of `r delay %>% count(destination_airport) %>% slice(which.max(n)) %>% pull(n)` and flights to `r delay %>% count(destination_airport) %>% slice(which.min(n)) %>% pull(destination_airport)` have the highest delay occurrences with a number of `r delay %>% count(destination_airport) %>% slice(which.min(n)) %>% pull(n)`.


<br>
```{r, echo = F}
cancel %>% 
  group_by(destination_airport) %>% 
  summarize(
    count = n()) %>% 
  mutate(
    destination_airport = fct_reorder(destination_airport, count),
    text_label = str_c("Airport: ", destination_airport, "\nCount: ", count)
      ) %>% 
  plot_ly(x = ~destination_airport, y = ~count, text = ~text_label, hoverinfo = "text",
          color = "rgb(255, 65, 54)", type = "bar", alpha = .5) %>% 
  layout(
    xaxis = list(title = "Destination Airport"),
    yaxis = list(title = "Count"),
    title = "Distribution of Cancellation by Destination Airport")
```



Flights from JFK to `r cancel %>% count(destination_airport) %>% slice(which.max(n)) %>% pull(destination_airport)` have the highest cancellation occurrences with a number of `r cancel %>% count(destination_airport) %>% slice(which.max(n)) %>% pull(n)` and flights to `r cancel %>% count(destination_airport) %>% slice(which.min(n)) %>% pull(destination_airport)` have the highest cancellation occurrences with a number of `r cancel %>% count(destination_airport) %>% slice(which.min(n)) %>% pull(n)`.

<br>
```{r}
delay_dest = function(dest){
  delay %>%
    filter(
      airline_name == dest
    ) %>% 
    group_by(destination_airport) %>% 
    summarize(
      count = n()
      ) %>% 
  mutate(
    destination_airport = fct_reorder(destination_airport, count),
    text_label = str_c("Airport: ", destination_airport, "\nCount: ", count)
      ) %>% 
  plot_ly(x = ~destination_airport, y = ~count, text = ~text_label, hoverinfo = "text", 
          color = ~destination_airport, type = "bar", alpha = .5) %>% 
  layout(
    xaxis = list(title = "Destination Airport"),
    yaxis = list(title = "Count"),
    title = "Distribution of Delay by Destination Airport")
}

cancel_dest = function(dest){
  cancel %>%
    filter(
      airline_name == dest
    ) %>% 
    group_by(destination_airport) %>% 
    summarize(
      count = n()
      ) %>% 
    mutate(
    destination_airport = fct_reorder(destination_airport, count),
    text_label = str_c("Airport: ", destination_airport, "\nCount: ", count)
      ) %>% 
    plot_ly(x = ~destination_airport, y = ~count, text = ~text_label, hoverinfo = "text",
          color = ~destination_airport, type = "bar", alpha = .5) %>% 
    layout(
      xaxis = list(title = "Destination Airport"),
      yaxis = list(title = "Count"),
      title = "Distribution of Cancellation by Destination Airport")
}
```

<br>


## Stratification by Airline {.tabset}

We can also take a look at whether different airlines could have different trends in delay and cancellation counts among all the destination airports.



### Alaska
```{r, echo = FALSE}
delay_dest("Alaska Airlines")
cancel_dest("Alaska Airlines")
```


### American
```{r, echo = FALSE}
delay_dest("American Airlines")
cancel_dest("American Airlines")
```

### Delta
```{r, echo = FALSE}
delay_dest("Delta Air Lines")
cancel_dest("Delta Air Lines")
```

### Endeavor
```{r, echo = FALSE}
delay_dest("Endeavor Air")
cancel_dest("Endeavor Air")
```

### JetBlue {.active}
```{r, echo = FALSE}
delay_dest("JetBlue Airways")
cancel_dest("JetBlue Airways")
```

### Republic
```{r, echo = FALSE}
delay_dest("Republic Airways")
cancel_dest("Republic Airways")
```

### United
```{r, echo = FALSE}
delay_dest("United Air Lines")
cancel_dest("United Air Lines")
```

<br>

## LAX and SFO {.tabset}
We found that LAX and SFO have outstanding delay and cancellation counts, so we decided to take a closer look at the underlying factors behind those delays and cancellations.

<br>

```{r, echo = F}
delay %>% 
  filter(
    destination_airport %in% c("LAX", "SFO"), 
    delay_minutes > 0
  ) %>% 
  plot_ly(x = ~delay_minutes, color = "rgb(255, 65, 54)", alpha = .5,
          type = "histogram") %>% 
    layout(
      xaxis = list(title = "Delay Time (minutes)", range = c(0, 300)),
      yaxis = list(title = "Count"), 
      title = "Distribution of Delay Time")
```

We can clearly observe that the delay times are clustered before 180 minutes. For the following explorations, we filtered the delay minutes between 0 to 180 minutes. 

<br>


### Delay by Airline
```{r, echo = F}
delay %>% 
  filter(
    destination_airport %in% c("LAX", "SFO"),
    delay_minutes %in% 0:180
  ) %>% 
  plot_ly(x = ~destination_airport, y = ~delay_minutes, color = ~airline_name,
          type = 'box', alpha = .5) %>% 
    layout(
      boxmode = "group",
      xaxis = list(title = "Destination Airport"),
      yaxis = list(title = "Delay Time (minutes)"),
      legend = list(title = list(text = '<b> Airline </b>')))
```
The airlines which departure from JFK to the two airports are different, and there is no distinct trend in delay time in minutes among different airlines between the two airports.  


### Delay by Month
```{r, echo = F}
delay %>% 
  filter(
    destination_airport %in% c("LAX", "SFO"),
    delay_minutes %in% 0:180
  ) %>% 
  mutate(
    month = recode(month, "11" = "Nov", "12" = "Dec", "1" = "Jan"),
    month = as.factor(month),
    month = fct_reorder(month, date)
  ) %>% 
  plot_ly(x = ~destination_airport, y = ~delay_minutes, color = ~month,
          type = 'box', alpha = .5) %>% 
    layout(
      boxmode = "group",
      xaxis = list(title = "Destination Airport"),
      yaxis = list(title = "Delay Time (minutes)"),
      legend = list(title = list(text = '<b> Month </b>')))
```
Both airports show an increasing trend in delay minutes from November to January. There is no distinct difference in delay time in minutes in different months between the two airports. 


### Cancellation by Hour
```{r, echo = F}

delay %>% 
  filter(
    destination_airport %in% c("LAX", "SFO")
  ) %>% 
  group_by(scheduled_hour, destination_airport) %>% 
  summarize(
    count = n()
  ) %>% 
  plot_ly(x = ~scheduled_hour, y = ~count, color = ~destination_airport,
          type = 'bar', alpha = .5) %>% 
    layout(
      xaxis = list(title = "Scheduled Hour"),
      yaxis = list(title = "Count"),
      legend = list(title = list(text = '<b> Airport </b>')))

```

We can observe a distinct difference in cancellation counts in each scheduled hour between the two airports.

<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>

