---
title: "Poisson model"
output: github_document
---

```{r}


library(dplyr)
library(tidyverse)
library(ggridges)
library(patchwork)
library(plotly)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))


cancel_raw = read_csv("tidied_data/cancel.csv")
covid = read_csv("tidied_data/covid.csv")
daily_weather = read_csv("tidied_data/daily_weather.csv")

```

# Preparations before modelling
## Outcome: mutate count cancellation data
Count numbers of cancellation per day
```{r}
cancel <- cancel_raw %>% 
  mutate(number = 1)%>% 
  mutate(number = as.numeric(number)) %>% 
  group_by(date) %>% 
  mutate(cancel_count = sum(number)) %>% 
  select(-flight_number,-destination_airport,-scheduled_hour,-scheduled_departure_time,-scheduled_elapsed_time_minutes, -number) %>%
  distinct

```

## Merge outcome and predictors:

weather dataset: select `daily_average_dry_bulb_temperature`,  `daily_average_relative_humidity`, `daily_peak_wind_speed` predictors
```{r}
weather <- daily_weather %>% 
  mutate(
    temperature = daily_average_dry_bulb_temperature,
    humidity = daily_average_relative_humidity,
    windspeed = daily_peak_wind_speed
         )  %>% 
  select(date, year, month, day, temperature, humidity, windspeed) 
  
  
```

merge dataset 
```{r}
# merge weather and covid dataset
weather_covid <- weather %>% left_join(covid, by = c("month", "day", "year")) %>% 
mutate(covid_case = case_count) %>% 
  select(-case_count)

```
## Final Tidied dataset
```{r}

# merge cancel and weather_covid dataset
cancel_tidy <- weather_covid %>% 
  left_join(cancel, by = c("month", "day", "year")) 

cancel_tidy <- cancel_tidy%>% 
  mutate(
    temperature = as.numeric(temperature),
    humidity = as.numeric(humidity),
    windspeed = as.numeric(windspeed),
    covid_case = as.numeric(covid_case),
    airline_name = as_factor(airline_name),
    month = ifelse(month == 11, "November", 
                        ifelse(month == 12, "December", "January")),
    month1 = ifelse(month == "November", "Nov", 
                        ifelse(month == "December", "Dec", "Jan")),
    year_month = paste(year, month1, sep="-")) %>% 
    filter(!is.na(cancel_count))  %>% 
    select(-month1, -date.y, -date.x)

write_csv(cancel_tidy, "tidied_data/cancel_tidy.csv")
 
```
# Check distribution
Since our outcome, `cancel_coun`t, is calculated using a count variable (number of cancellation) repeated over time, we’ll model it using a Poisson regression model.
```{r}
# poisson distribution of counts
cancel_tidy %>% 
  ggplot(aes(x = cancel_count)) + 
  geom_histogram(binwidth = 1.8) +
  theme_light() +
  labs(x = "Numbers of cancellation per day",
       y = "Count",
       title = "Distribution of cancellation counts per day, 2021/11/1 to 2022/1/31") +
  theme(plot.title = element_text(hjust = 0.5))
```

## Stratification factors
Cancellation count per month show difference both by month and airline, therefore, using month and airline as stratification factors
```{r}
cancel_airline <- cancel_tidy  %>%
  group_by(year_month, airline_name) %>%
  mutate(Total_number_of_cancellation = sum(cancel_count)) 

cancel_airline %>%
  select(year_month, airline_name, Total_number_of_cancellation) %>%
  distinct %>% 
  pivot_wider(
  names_from = airline_name, 
  values_from = Total_number_of_cancellation) %>%
  head() %>% 
  knitr::kable(digits = 2, caption = "Total number of cancellation in each airline") 

write_csv(cancel_airline, "tidied_data/cancel_airline.csv")
```

```{r}
plot_cancel_airline <- cancel_airline %>%
  ggplot(aes(x = year_month, y = Total_number_of_cancellation, fill = airline_name)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  labs(
    title = "Total Number of Cancellation by Airline",
    x = "Year and Month",
    y = "Total Number of Cancellation"
  ) +
  theme(legend.position="right", legend.title = element_blank(),
        text = element_text(size = 10),
        axis.text.x = element_text(angle = 60, hjust = 1, size = 5)) +
  facet_grid(. ~ airline_name)

  
ggplotly(plot_cancel_airline)
```

# Poisson model
```{r}
poisson = glm(cancel_count ~ temperature + humidity + windspeed + covid_case,family = "poisson",data=cancel_tidy)


summary(poisson) 
```

# Poisson model Nested by month
```{r}
cancel_tidy %>% 
  nest(df = -month) %>%
  mutate(
    models = map(.x = df, ~ glm(cancel_count ~ temperature + humidity + windspeed + covid_case,family = "poisson", data = .x)),
    results = map(models, broom::tidy)
  ) %>% 
  unnest(results) %>% 
  select(month, term, estimate) %>% 
  mutate(term = fct_inorder(term)) %>% 
  pivot_wider(
    names_from = term, 
    values_from = estimate) %>% 
  knitr::kable(digits = 6, caption = "Poisson model nested by month")
```

## OR and 95% CI
```{r}
poisson_by_month = cancel_tidy %>%
  nest(data = -month) %>% 
  mutate(
    models = map(.x = data, ~glm(cancel_count ~ temperature + humidity + windspeed + covid_case, family = "poisson", data = .x)),
    results = map(models, broom::tidy)
    ) %>% 
  select(month, results) %>% 
  unnest(results) %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error),
    p.value = format(p.value, scientific = TRUE, digits = 3)
  ) %>% 
  select(month, term, OR, CI_lower,CI_upper, p.value) 

poisson_by_month %>% 
  filter(term != "(Intercept)" ) %>% 
  knitr::kable(digits = 3, align = "llccc", col.names = c("Month", "Terms", "Estimated adjusted OR", "CI lower bound", "CI upper bound", "P-value"))
```

## Plot
Create a plot showing the estimated ORs and CIs for each month

```{r}
poisson_by_month %>%
  filter(term != "(Intercept)") %>%
  ggplot(aes(x = month, y = OR, color = term)) + 
  geom_point(show.legend = FALSE, aes()) +
  geom_errorbar(aes(ymin = CI_lower, 
                    ymax = CI_upper)) +
  labs(
    title = "Estimated OR with 95% CI in Cancellation Count Data by Month",
    x = "Month",
    y = "Estimated OR with CI"
  ) +
  theme(legend.position="right", legend.title = element_blank(),
        text = element_text(size = 10),
        axis.text.x = element_text(angle = 60, hjust = 1, size = 8)) + 
  facet_grid(. ~ term)

```


# Poisson model nested by airline


```{r}
cancel_tidy %>% 
  nest(df = -airline_name) %>%
  mutate(
    models = map(.x = df, ~ glm(cancel_count ~ temperature + humidity + windspeed + covid_case,family = "poisson", data = .x)),
    results = map(models, broom::tidy)
  ) %>% 
  unnest(results) %>% 
  select(airline_name, term, estimate) %>% 
  mutate(term = fct_inorder(term)) %>% 
  pivot_wider(
    names_from = term, 
    values_from = estimate) %>% 
  knitr::kable(digits = 6, caption = "Poisson model nested by airline")
```
## OR and 95% CI
```{r}
  
poisson_by_airline = cancel_tidy %>%
  nest(data = -airline_name) %>% 
  mutate(
    models = map(.x = data, ~glm(cancel_count ~ temperature + humidity + windspeed + covid_case, family = "poisson", data = .x)),
    results = map(models, broom::tidy)
    ) %>% 
  select(airline_name, results) %>% 
  unnest(results) %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error),
    p.value = format(p.value, scientific = TRUE, digits = 3)
  ) %>% 
  select(airline_name, term, OR, CI_lower,CI_upper, p.value) 

poisson_by_airline %>% 
  filter(term != "(Intercept)" ) %>% 
  knitr::kable(digits = 3, align = "llccc", col.names = c("Airline Name", "Terms", "Estimated adjusted OR", "CI lower bound", "CI upper bound", "P-value"))
```

## Plot
Create a plot showing the estimated ORs and CIs for each airline

```{r}
poisson_by_airline %>%
  filter(term != "(Intercept)") %>%
  ggplot(aes(x = airline_name, y = OR, color = term)) + 
  geom_point(show.legend = FALSE, aes()) +
  geom_errorbar(aes(ymin = CI_lower, 
                    ymax = CI_upper)) +
  labs(
    title = "Estimated OR with 95% CI in Cancellation Count Data by Airline",
    x = "Airline",
    y = "Estimated OR with CI"
  ) +
  theme(legend.position="right", legend.title = element_blank(),
        text = element_text(size = 10),
        axis.text.x = element_text(angle = 60, hjust = 1, size = 8)) + 
  facet_grid(. ~ term)

```

