---
title: "Regression_delay"
author: "Fengyi Ma"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(modelr)
library(mgcv)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
set.seed(1)
```

## Data Import
```{r}
delay = read.csv("./tidied_data/delay.csv") %>% 
  janitor::clean_names() 

h_weather = read.csv("./tidied_data/hourly_weather.csv") %>% 
  janitor::clean_names() 
```

```{r}

```

### hypothesized predictors
At 100% relative humidity, the wet-bulb temperature is equal to the air temperature (dry-bulb temperature); at lower humidity the wet-bulb temperature is lower than dry-bulb temperature because of evaporative cooling.

The wet-bulb temperature is defined as the temperature of a parcel of air cooled to saturation (100% relative humidity) by the evaporation of water into it, with the latent heat supplied by the parcel.
