---
title: "Regression_delay"
author: "Fengyi Ma"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(modelr)
library(mgcv)
library(dplyr)
library(tidyr)
library(skimr)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
set.seed(1)
```

## Data Import

```{r}
delay = read.csv("./tidied_data/delay.csv") %>% 
  janitor::clean_names() 

h_weather = read.csv("./tidied_data/hourly_weather.csv") %>% 
  janitor::clean_names() 
```

## clean dataset 'delay'

Check how many airlines
```{r}
unique(delay$airline_name)
```
_7_ -> ok, keep

Check how many destinations
```{r}
unique(delay$destination_airport)
```
_66_ -> too many, remove

Keep variables of interest and `date` for merge purpose (which will be removed later) 
```{r}
delay = delay %>% 
  mutate(
    airline = airline_name,
    hour = scheduled_hour,
    delay = delay_minutes,
    carrierd = delay_carrier_minutes,
    extrmwd = delay_weather_minutes,
    nasd = delay_national_aviation_system_minutes,
    securityd = delay_security_minutes,
    latarrd = delay_late_aircraft_arrival_minutes) %>% 
  mutate(hour = as.numeric(hour)) %>% 
  select(airline, date, month, hour, delay, carrierd, extrmwd, nasd, securityd, latarrd)
```

check 'NA'
```{r eval=FALSE}
sum(is.na(delay))
```
_0_ -> good


## clean dataset 'h_weather'

About the measure of temperature:
Since the dry bulb temperature is the ambient air temperature measured by regular thermometers, that is, the temperature often mentioned in our general weather forecast. Thus, we decide to use the variable `hourly_dry_bulb_temperature` to represent temperature.

```{r}
h_weather = h_weather %>% 
  mutate(
    temperature = hourly_dry_bulb_temperature,
    humidity = hourly_relative_humidity,
    visibility = hourly_visibility,
    wind_s = hourly_wind_speed,
    hour = as.numeric(hour)) %>% 
      select(date, month, hour, temperature, humidity, visibility, wind_s)
```

check 'NA'
```{r eval=FALSE}
sum(is.na(h_weather))
```
_0_ -> good


## merge datasets 'delay' and 'hourly_weather'
```{r}
raw_df = merge(x = delay, y = h_weather, by = c("date", "month", "hour"),
               all.x = TRUE)
```

check 'NA'
```{r eval=FALSE}
sum(is.na(raw_df))
```
_0_ -> good

To this step, we have our raw dataset for doing association analysis.
Since the outcome variable `delay` is a continuous variable, we would do linear regression.
However, there might be too many variables so far (ignoring `date` we still got 12 potential predictors), so next step will be fitting the model.

By intuition, I would set:

`airline` -> categorical
`month` -> categorical
`hour` -> categorical (need further categorization)
and for the rest -> continuous

Let's first check how each variable roughly distribute.

### For categorical variables
```{r}
cat_sum = raw_df %>% 
  select(airline, month, hour) %>% 
  mutate(
    airline = as.factor(airline),
    month = as.factor(month),
    hour = as.factor(hour)
    ) %>% 
  summary(maxsum = 24)

cat_sum
```

### For continuous variables
```{r}
con_sum_df = raw_df %>% 
  select(-date, -airline, -month, -hour)

con_sum = skim(con_sum_df) %>%
  dplyr::select(-n_missing, -complete_rate) %>% 
  mutate(
    mean = numeric.mean,
    sd = numeric.sd,
    histogram = numeric.hist,
    var = skim_variable,
    min = numeric.p0,
    max = numeric.p100,
    median = numeric.p50,
    q1 = numeric.p25,
    q3 = numeric.p75
  ) %>% 
  dplyr::select(-numeric.mean, -numeric.sd, -numeric.hist, -skim_variable, -numeric.p0, -numeric.p100, -numeric.p50, -numeric.p25, -numeric.p75) %>% 
  relocate(var, mean, min, q1, median, q3, max, sd, histogram)

con_sum
```

Next, as mentioned above, we want to further categorize variable `hour`.

_Motivation:_ 

To increase the power of our model by reducing the number of parameters involved and to be more efficient and concise.

_Rationale for the categorization of `hour`:_

Based on the previous inspection of variable `hour`, we could see that: except for 5, 8, 10, 22, and 23, the frequencies of the other classes are roughly even (between 1000-2000). Take this into consideration, our rationale for classification will be a combination of convention and the desire to achieve a uniform distribution.

Thus, we would categorize `hour` into the following 4 categories:

_morning_: 5-8

_noon_: 9-13

_afternoon_: 14-17 

_night_:18-23 

Now, we can start categorize `hour`
```{r}
raw_df = raw_df %>% 
  mutate(hour_c = cut(hour, breaks = c(4, 8, 13, 17, 24),
                      labels = c("morning","noon","afternoon","night"))) %>% 
  select(-hour)
```

As usual, check if it was done properly
```{r}
summary(as.factor(raw_df$hour_c))
sum(is.na(raw_df))
```

_0_ 'NA' and the distribution looks good.

Now, since we are not yet able to decide the variable type of the rest of the predictors, we decide to do univariate linear regression for each variable first.

## Univariate linear regression

### Continuous vars

* temperature

```{r message=FALSE, warning=FALSE}
lrTemp = lm(delay~temperature, data = raw_df)
summary(lrTemp) %>% broom::glance()
summary(lrTemp) %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value)
```

* humidity

```{r message=FALSE, warning=FALSE}
lrHum = lm(delay~humidity, data = raw_df)
summary(lrHum) %>% broom::glance()
summary(lrHum) %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value)
```

* visibility

```{r message=FALSE, warning=FALSE}
lrVis = lm(delay~visibility, data = raw_df)
summary(lrVis) %>% broom::glance()
summary(lrVis) %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value)
```

* wind speed

```{r message=FALSE, warning=FALSE}
lrWin = lm(delay~wind_s, data = raw_df)
summary(lrWin) %>% broom::glance()
summary(lrWin) %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value)
```

* carrier delay

```{r message=FALSE, warning=FALSE}
lrCar = lm(delay~carrierd, data = raw_df)
summary(lrCar) %>% broom::glance()
summary(lrCar) %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value)
```

* extreme weather delay

```{r message=FALSE, warning=FALSE}
lrExw = lm(delay~extrmwd, data = raw_df)
summary(lrExw) %>% broom::glance()
summary(lrExw) %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value)
```

* NAS delay

```{r message=FALSE, warning=FALSE}
lrNas = lm(delay~nasd, data = raw_df)
summary(lrNas) %>% broom::glance()
summary(lrNas) %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value)
```

* security delay

```{r message=FALSE, warning=FALSE}
lrSec = lm(delay~securityd, data = raw_df)
summary(lrSec) %>% broom::glance()
summary(lrSec) %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) 
```

* late arrival delay

```{r message=FALSE, warning=FALSE}
lrLat = lm(delay~latarrd, data = raw_df)
summary(lrLat) %>% broom::glance()
summary(lrLat) %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value)
```

### Categorical vars

First, tell R that they are categorical variables
```{r}
raw_df = 
  raw_df %>% 
  mutate(
    month = fct_infreq(as.factor(month)),
    hour_c = fct_infreq(hour_c),
    airline = fct_infreq(airline)
    )
```

* month

```{r message=FALSE, warning=FALSE}
lrMon = lm(delay~month, data = raw_df)
summary(lrMon) %>% broom::glance()
summary(lrMon) %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value)
```

* hour

```{r}
lrHour = lm(delay~hour_c, data = raw_df)
summary(lrHour) %>% broom::glance()
summary(lrHour) %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value)
```

* airline

```{r}
lrAL = lm(delay~airline, data = raw_df)
summary(lrAL) %>% broom::glance()
summary(lrAL) %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value)
```

The above results were not very satisfying, however, according to the R-squared values, we would choose to include `carrierd` and `latarrd` in to the final model.



