---
title: "Poisson model"
output: 
  html_document:
    toc: true
    toc_float: true
    output: 
  always_allow_html: true

---

```{r}


library(dplyr)
library(tidyverse)
library(ggridges)
library(patchwork)
library(plotly)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

cancel_airline = read_csv("tidied_data/cancel_airline.csv")
cancel_tidy = read_csv("tidied_data/cancel_tidy.csv")

```
# Overview
We wanted to investigate factors that are related to - and may be used to predict daily flight cancellation count since the effects of weather and COVID-19 were our primary research interests, but we did not have any precise hypothesis. For instance, there can be daily flight cancellations because of COVID breakouts or lockdowns. 

# Data description
In this research, we look for a relationship between cancellation and a number of variables. The number of cancellations each day is therefore regarded as the result (dependent variable), and exposures of interest are regarded as the independent variables.

`cancel_coun`t: daily flight cancellation count in JFK airport
`temperature`: daily average dry bulb temperature
`humidity`: daily average relative humidity
`windspeed`: daily average relative windspeed
`covid_case`: new daily cases of COVID-19
`airline_name`: domestic airline 
`year_month`: the date format in "Year Month"
 
# Pre-analysis
## Check distribution
Since our outcome, `cancel_count`, is calculated using a count variable (number of cancellation) repeated over time, we’ll model it using a Poisson regression model.
```{r}
# poisson distribution of counts
his_plot<- cancel_tidy %>% 
  plot_ly(x = ~cancel_count, 
   type = "histogram", marker = list(color = 'rgb(255, 156, 128)'))

his_plot %>% 
         layout(
         title= 'Distribution of daily flight cancellation, 2021/11/1 to 2022/1/31',
         xaxis = list(title = "Daily flight cancellation", tickangle = -45),
         yaxis = list(title = "Count"))

```

## Stratification factors
Cancellation count per month show difference both by month and airline, therefore, using month and airline as stratification factors.
```{r}
cancel_airline <- cancel_tidy  %>%
  group_by(year_month, airline_name) %>%
  mutate(Total_number_of_cancellation = sum(cancel_count)) %>%   
  select(year_month, airline_name, Total_number_of_cancellation) %>%
  distinct %>%
  mutate(year_month = fct_relevel(year_month, "2021-Nov", "2021-Dec", "2022-Jan"))  

cancel_airline %>%
  pivot_wider(
  names_from = airline_name, 
  values_from = Total_number_of_cancellation) %>%
  head() %>% 
  knitr::kable(digits = 2, caption = "Monthly flight Cancellation numbers by month and airline") 

```


```{r}
bar_plot1 = cancel_airline %>% 
  mutate(year_month = fct_relevel(year_month, "2021-Nov", "2021-Dec", "2022-Jan")) %>% 
  plot_ly(x = ~year_month, y = ~Total_number_of_cancellation, color = ~airline_name, type = "bar", colors = "OrRd")

bar_plot1 = bar_plot1 %>% 
  layout(xaxis = list(title = "Time"),
         yaxis = list(title = "Monthly flight cancellation numbers"),
         margin = list(b = 100),
         barmode = 'group')
bar_plot1

bar_plot2 = cancel_airline %>% 
  mutate(year_month = fct_relevel(year_month, "2021-Nov", "2021-Dec", "2022-Jan")) %>% 
  plot_ly(x = ~airline_name, y = ~Total_number_of_cancellation, color = ~year_month, type = "bar", colors = "OrRd")

bar_plot2 = bar_plot2 %>% 
  layout(xaxis = list(title = "Airline Name", tickangle = -45),
         yaxis = list(title = "Monthly flight cancellation numbers"),
         margin = list(b = 100),
         barmode = 'group')
bar_plot2
```
lag covid-19
```{r}
cancel_covid <- cancel_tidy  %>%
  select(date, cancel_count, covid_case) %>% 
  distinct()

cancel_plot <-  cancel_covid  %>%
  mutate(text_label = str_c("Date: ", date, "\nCancellation flight: ", cancel_count)) %>% 
  plot_ly(
    x = ~date, y = ~cancel_count, type = "scatter", mode = "lines+markers",
   text = ~text_label, alpha = 0.5, name = 'Daily flight cancellation') %>% 
  layout(
         yaxis = list(title = "Daily flight cancellation"),
         annotations = list( 
    list(x = 0.1 , y = 0.8, text = "Daily flight cancellation", showarrow = F, xref='paper', yref='paper')))
         
covid_plot <-  cancel_covid  %>%
  mutate(text_label = str_c("Date: ", date, "\nCovid cases: ", covid_case)) %>% 
  plot_ly(
    x = ~date, y = ~covid_case, type = "scatter", mode = "lines+markers",
   text = ~text_label, alpha = 0.5, name = 'Daily covid cases') %>% 
  layout(
         yaxis = list(title = "Daily covid cases"),
         annotations = list( 
    list(x = 0.1 , y = 0.8, text = "Daily covid cases", showarrow = F, xref='paper', yref='paper')))

scatter_plot <- subplot(cancel_plot, covid_plot, nrows = 2) %>% 
  layout(title = list(text = "Time trend of daily cancellation vs. daily covid cases"),
         plot_bgcolor='#e5ecf6') 
scatter_plot


```

# Poisson model
```{r}
poisson = glm(cancel_count ~ temperature + humidity + windspeed + covid_case,family = "poisson",data=cancel_tidy)


summary(poisson) 
```

# Poisson model Nested by month
```{r}
cancel_tidy %>% 
  nest(df = -month) %>%
  mutate(
    models = map(.x = df, ~ glm(cancel_count ~ temperature + humidity + windspeed + covid_case,family = "poisson", data = .x)),
    results = map(models, broom::tidy)
  ) %>% 
  unnest(results) %>% 
  select(month, term, estimate) %>% 
  mutate(term = fct_inorder(term)) %>% 
  pivot_wider(
    names_from = term, 
    values_from = estimate) %>% 
  knitr::kable(digits = 6, caption = "Poisson model nested by month")
```

## OR and 95% CI
```{r}
poisson_by_month = cancel_tidy %>%
  nest(data = -month) %>% 
  mutate(
    models = map(.x = data, ~glm(cancel_count ~ temperature + humidity + windspeed + covid_case, family = "poisson", data = .x)),
    results = map(models, broom::tidy)
    ) %>% 
  select(month, results) %>% 
  unnest(results) %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error),
    p.value = format(p.value, scientific = TRUE, digits = 3)
  ) %>% 
  select(month, term, OR, CI_lower,CI_upper, p.value) 

poisson_by_month %>% 
  filter(term != "(Intercept)" ) %>% 
  knitr::kable(digits = 3, align = "llccc", col.names = c("Month", "Terms", "Estimated adjusted OR", "CI lower bound", "CI upper bound", "P-value"))
```

## Plot
Create a plot showing the estimated ORs and CIs for each month

```{r}
poisson_by_month %>%
  filter(term != "(Intercept)") %>%
  ggplot(aes(x = month, y = OR, color = term)) + 
  geom_point(show.legend = FALSE, aes()) +
  geom_errorbar(aes(ymin = CI_lower, 
                    ymax = CI_upper)) +
  labs(
    title = "Estimated OR with 95% CI in Cancellation Count Data by Month",
    x = "Month",
    y = "Estimated OR with CI"
  ) +
  theme(legend.position="right", legend.title = element_blank(),
        text = element_text(size = 10),
        axis.text.x = element_text(angle = 60, hjust = 1, size = 8)) + 
  facet_grid(. ~ term)

```


# Poisson model nested by airline


```{r}
cancel_tidy %>% 
  nest(df = -airline_name) %>%
  mutate(
    models = map(.x = df, ~ glm(cancel_count ~ temperature + humidity + windspeed + covid_case,family = "poisson", data = .x)),
    results = map(models, broom::tidy)
  ) %>% 
  unnest(results) %>% 
  select(airline_name, term, estimate) %>% 
  mutate(term = fct_inorder(term)) %>% 
  pivot_wider(
    names_from = term, 
    values_from = estimate) %>% 
  knitr::kable(digits = 6, caption = "Poisson model nested by airline")
```
## OR and 95% CI
```{r}
  
poisson_by_airline = cancel_tidy %>%
  nest(data = -airline_name) %>% 
  mutate(
    models = map(.x = data, ~glm(cancel_count ~ temperature + humidity + windspeed + covid_case, family = "poisson", data = .x)),
    results = map(models, broom::tidy)
    ) %>% 
  select(airline_name, results) %>% 
  unnest(results) %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error),
    p.value = format(p.value, scientific = TRUE, digits = 3)
  ) %>% 
  select(airline_name, term, OR, CI_lower,CI_upper, p.value) 

poisson_by_airline %>% 
  filter(term != "(Intercept)" ) %>% 
  knitr::kable(digits = 3, align = "llccc", col.names = c("Airline Name", "Terms", "Estimated adjusted OR", "CI lower bound", "CI upper bound", "P-value"))
```

## Plot
Create a plot showing the estimated ORs and CIs for each airline

```{r}
poisson_by_airline %>%
  filter(term != "(Intercept)") %>%
  ggplot(aes(x = airline_name, y = OR, color = term)) + 
  geom_point(show.legend = FALSE, aes()) +
  geom_errorbar(aes(ymin = CI_lower, 
                    ymax = CI_upper)) +
  labs(
    title = "Estimated OR with 95% CI in Cancellation Count Data by Airline",
    x = "Airline",
    y = "Estimated OR with CI"
  ) +
  theme(legend.position="right", legend.title = element_blank(),
        text = element_text(size = 10),
        axis.text.x = element_text(angle = 60, hjust = 1, size = 8)) + 
  facet_grid(. ~ term)

```

